Supplementary Materials

### 2022-12-08 manuscript/supplementary.tex
ms changes

Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity is calculated as

\begin{align}
  \bar{\pi}(l) = \frac{n_D}{n_T}
\end{align}

@@

\begin{align}
  \bar{\pi}(l) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
\end{align}

@@

\begin{align*}
  \bar{\pi}(\mathcal{A}) = \frac{\sum_{l \in \mathcal{A}} \bar{\pi}(l)}{L_\mathcal{A}}.
\end{align*}

where $L_\mathcal{A} = |\mathcal{A}|$ is the number of accessible bases. Note
that both $L_\mathcal{A}$ and the sum are random variables in genomic data.
The denominator varies along the genome, based on the number of
reliably-genotyped and accessible bases in the reference genome. This is a
ratio estimator, and is thus biased, since by Jensen's inequality
$\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$.

We explore this with block-bootstrap bias correction. TODO

>>>>

Thus $\mathbf{P}_l = [n_S, n_D]$, and the pairwise diversity across the $n$
sampled chromosomes is calculated as

\begin{align}
  \pi(l) = \frac{n_D}{n_T}
\end{align}

@@

\begin{align}
  \pi(l) = \frac{2}{n(n-1)} \sum_{i < j}^n k_{i,j}
\end{align}

@@

\begin{align*}
  \bar{\pi}(\mathcal{A}) = \frac{\sum_{l \in \mathcal{A}} \pi(l)}{L_\mathcal{A}}.
\end{align*}

where $L_\mathcal{A} = |\mathcal{A}|$ is the number of accessible bases. Note
that the sum in the numerator is random over the sample of chromosomes sampled
from the population. Estimates of pairwise diversity often condition on the
accessible bases, and thus treat this as fixed. However, the number of
accessible bases varies across the chromosome; this can lead to a source of
apparent bias during block-bootstrap estimates of uncertainty.  In this case,
pairwise diversity is a ratio estimator, and is thus biased, since by Jensen's
inequality $\E(\nicefrac{y}{x}) \ge \nicefrac{\E(y)}{\E(x)}$ for random
variables $x$ and $y$.
